---
title: "Hands-on Exercise 1"
author: "Jaya George"
---

# *Getting Started*

## *Installing and loading the required libraries*

```{r}
library("pacman")

```

```{r}
pacman::p_load(tidyverse)
```

```{r}
library(readr)
```

## *Importing data*

```{r}
exam_data <- read_csv("C:/jayageorge7/ISS608-VAA/Hands-on_Ex/Hands-on_Ex01/data/Exam_data.csv",show_col_types = FALSE)
```
# *Introducing ggplot*

## *R Graphics VS ggplot*

### *R Graphics*

```{r}
hist(exam_data$MATHS)

```

```{r}
library(ggplot2)
```

### *ggplot2*

```{r}
ggplot(data=exam_data, aes(x = MATHS)) + geom_histogram(bins=10, boundary = 100, color="black", fill="grey") + ggtitle("Distribution of Maths scores")
```

# *Grammar of graphics*


## *Essential Grammatical Elements in ggplot2: data*

```{r}
ggplot(data=exam_data)
```

## *Aesthetic mappings*

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

## *geom*

### *geom_bar*

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

### *geom_dotplot*

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

### *using scale_y_continuous*

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

### *geom_histogram()*

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()       
```

### *geom_histogram() with fill*

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

### *geom_histogram() with fill by sub-group*

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

### *geom-density()*

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
```

### *geom-density() with fill by sub-group*

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

### *geom_boxplot()*

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            
```

### *geom_boxplot() in notched plot*

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

### *geom_violin()*

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

### *geom_point()*

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()            
```

### *Combination of geom objects*

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             size = 0.5)        
```

### **geom_boxplot()**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot()
```

## **Exploring statistical functions**

###**using stat_summary()**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun="mean",         
               colour ="red",        
               size=4)               
```

### **using geom() method**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun="mean",           
             colour ="red",          
             size=4)          
```

### **plot best fit curve**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(linewidth=0.5)
```

### **override default smoothing method**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5)
```

## **Essential Grammatical Elements in ggplot2: Facets**

### **Working with `facet_wrap()**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

### **facet_grid() function**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```

## **Essential Grammatical Elements in ggplot2: Coordinates**

### **Working with Coordinate**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

### using coord_flip

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

### **Changing the y- and x-axis range**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5)
```

## **Essential Grammatical Elements in ggplot2: themes**

### **Working with theme**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
```

### **using theme_classic()**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

### **using theme_minimal()**  

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```
