---
title: "Take-home Exercise 1"
author: "Jaya George"
date: "April 17,2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Overview

### Setting the scene

There are two major residential property market in Singapore, namely public and private housing. Public housing aims to meet the basic need of the general public with monthly household income less than or equal to S\$14,000. For families with monthly household income more than S\$14,000, they need to turn to the private residential market.

### Our task

In this exercise, Exploratory Data Analysis (EDA) methods and ggplot functions are used to explore:

-   the private residential markets and sub-markets of Singapore

-   the variations of unit price,unitsize etc between different property types,sale types

## Getting started

### Load packages

We load the following R packages using the `pacman::p_load()` function:

-   **tidyverse**: Core collection of R packages designed for data science

-   **haven**: To read in data formats such as SAS and SPSS

-   **ggrepel**: to provides geoms for **ggplot2** to repel overlapping text labels

-   **ggthemes**: to use additional themes for **ggplot2**

-   **patchwork**: to prepare composite figure created using **ggplot2**

-   **ggridges**: to plot ridgeline plots

-   **ggdist**: for visualizations of distributions and uncertainty

-   **scales**: provides the internal scaling infrastructure used by **ggplot2**

```{r}
library("pacman")
suppressWarnings(("lubridate"))
library(readr)
```

```{r}
pacman::p_load(tidyverse, haven,
               ggrepel, ggthemes,
               ggridges, ggdist,colorspace,ggdist,
               patchwork, scales)
```

### Import data

The dataset (*data/*ResidentialTransaction20240414220633.csv) used in the exercise is a subset of the transaction data of [REALIS](https://www.ura.gov.sg/reis/index) questionnaire data file retrieved from the <https://www.ura.gov.sg/reis/index>. It has been filtered to include only the private residential transaction data for q1 2024.

We import this dataset as res_data.

```{r}
#res_data1 <- read_csv("data/ResidentialTransaction20240308160536.csv",show_col_types = FALSE)
#res_data2 <- read_csv("data/ResidentialTransaction20240308160736.csv",show_col_types = FALSE)
#res_data3 <- read_csv("data/ResidentialTransaction20240308161009.csv",show_col_types = FALSE)
#res_data4 <- read_csv("data/ResidentialTransaction20240308161109.csv",show_col_types = FALSE)
#res_data5 <- read_csv("data/ResidentialTransaction20240414220633.csv",show_col_types = FALSE)
#res_data <- bind_rows(res_data1, res_data2, res_data3, res_data4, res_data5)
```

```{r}
res_data <- read_csv("data/ResidentialTransaction20240414220633.csv",show_col_types = FALSE)
```

```{r}
nrow(res_data)
```

```{r}
#res_data_q1_2024 <- res_data_final %>%
 # mutate(Sale_Date = as.Date(Sale_Date, format = "%d %b %Y")) %>% 
  #filter(quarter(Sale_Date) == 1 & year(Sale_Date) == 2024)
```

### Data pre-processing

We first take a look at the data, and check if there are any duplicate entries.

```{r}
glimpse(res_data)
```

#### **Check for duplicates**

Using the `duplicated` function, we see that there are no duplicate entries in the data.

```{r}
res_data[duplicated(res_data),]
```

#### **Check and handle missing values**

```{r}
for(column_name in names(res_data)) {
  na_count <- sum(is.na(res_data[[column_name]]))

  if (na_count > 0) {
    message("Column '", column_name, "' has ", na_count, " NA values.")
  }
}
```

#### **Filtering data for selected variables**

Next, we want to reduce the size of the dataset to focus on the variables that would be suitable for this exercise. Primarily referred to the <https://www.ura.gov.sg/reis/dataDictionary>

The primary columns/variables that will be used in this exercise are:

+-----------------------+----------------------------------------------------------------------------------------------+
| Field Name            | Field Description                                                                            |
+=======================+==============================================================================================+
| Property Type         | Terrace House,Semi-Detached House,Executive Condominium,Detached House,Condominium,Apartment |
+-----------------------+----------------------------------------------------------------------------------------------+
| Area (SQFT)           | Area of the property in sqft                                                                 |
+-----------------------+----------------------------------------------------------------------------------------------+
| Unit Price (\$ PSF)   | Unit Price of the property in sqft                                                           |
+-----------------------+----------------------------------------------------------------------------------------------+
| Sale Date             | Date of the sale transaction                                                                 |
+-----------------------+----------------------------------------------------------------------------------------------+
| Type of Sale          | New Sale,Resale,Sub Sale                                                                     |
+-----------------------+----------------------------------------------------------------------------------------------+
| Planning Region       | Central,East,West,North,North East                                                           |
+-----------------------+----------------------------------------------------------------------------------------------+
| Planning Area         |                                                                                              |
+-----------------------+----------------------------------------------------------------------------------------------+

```{r}
res_data_q1_2024 = res_data %>% 
  select("Property Type", "Area (SQFT)","Unit Price ($ PSF)","Sale Date","Transacted Price ($)","Postal District","Type of Sale","Planning Region","Planning Area","Purchaser Address Indicator") %>%
  rename("Property_Type" = "Property Type",
         "Area_sqft" = "Area (SQFT)",
         "Unit_Price_sqft"="Unit Price ($ PSF)",
         "Sale_Date"="Sale Date",
         "Sale_Price"="Transacted Price ($)",
         "Postal_District"="Postal District",
         "Sale_Type"="Type of Sale",
         "Planning_Region"="Planning Region",
         "Planning_Area"="Planning Area",
         "Purchaser_Address_Indicator"="Purchaser Address Indicator")
```

## EDA 1: Distribution of transaction counts across Planning Reions and Property Types

To visualize the transaction counts among different property types in different Planning regions and drill down to planning area.

We plot a stacked bar chart that shows the transaction counts across the planning regions and the stacks shows the split of counts between different Property Types.

::: panel-tabset
## Plot

```{r, fig.width=12, fig.height=10}
#| echo: false

transaction_counts <- res_data_q1_2024 %>%
  group_by(Planning_Area) %>%  # Group by planning area
  count(Property_Type) %>%     # Count transactions per property type within each area
  mutate(total_transactions = sum(n))  # Add a new column with total transactions per area
 

p2<- ggplot(transaction_counts, aes(x = Planning_Area, y = n, fill = Property_Type)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Transaction Count by Property Type and Sale Type (Q1 2024)",
    x = "Property Type",
    y = "Number of Transactions"
  ) +
  theme(legend.position = "none",axis.text.x = element_text(angle = 90, hjust = 1))


transaction_counts <- res_data_q1_2024 %>%
  count(Planning_Region,Property_Type) 

p1<- ggplot(transaction_counts, aes(x = Planning_Region, y = n, fill = Property_Type)) +
  geom_bar(stat = "identity", position = "stack") + 
  labs(
    title = "Transaction Count by Property Type and Sale Type (Q1 2024)",
    x = "Property Type",
    y = "Number of Transactions"
  ) +
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 45, hjust = 1)) 


(p1/p2)
```

## Code

```{r}
#| eval: false

#abc
```
:::

+----------------------------------------------------------------------------------------------------+
| InSights                                                                                           |
+====================================================================================================+
| -   Overall sale of apartments seem to be the highest amongst all property types                   |
|                                                                                                    |
| -   Apartment sale in Central Region seems to be highest amongst all regions                       |
|                                                                                                    |
| -   Drilling down to Planning Area, apartment sale seems to be highest in Ang mo kio planning area |
+----------------------------------------------------------------------------------------------------+

## EDA 2: Distribution of unit size and unit price per sqft across property types

::: panel-tabset
## Plot

```{r, fig.width=12, fig.height=8}
#| echo: false
# Create the ridge plot
ridgeplot_size <- ggplot(res_data_q1_2024, aes(x = Area_sqft, y = Property_Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  #facet_wrap(~ Planning_Region, scales = "free_x") + # Faceting by subject
  labs(title = "                                                            Property Types: Size and Price Range", hjust = 0.5,
       subtitle = "Density plots of Area_sqft stratified by Property_Type",
       x = "Area_sqft") +
  scale_x_continuous(breaks = seq(0, max(res_data_q1_2024$Area_sqft), by = 1000))+
                       #limits = c(0, 8000))+
  theme_minimal() +
  theme(legend.position = "none", # Hides the legend,
        plot.title = element_text(face = "bold", size = 12),
        plot.subtitle = element_text(size = 8))


ridgeplot_size_unit_price <- ggplot(res_data_q1_2024, aes(x = Unit_Price_sqft, y = Property_Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  #facet_wrap(~ Planning_Region, scales = "free_x") + # Faceting by subject
  labs(subtitle = "Density plots of Unit_Price_sqft stratified by Property_Type",
       x = "Unit_Price_sqft") +
  scale_x_continuous(breaks = seq(0, max(res_data_q1_2024$Unit_Price_sqft), by = 500))+
  theme_minimal() +
  theme(legend.position = "bottom", # Hides the legend,
        plot.title = element_text(face = "bold", size = 12),
        plot.subtitle = element_text(size = 8))


(ridgeplot_size/ridgeplot_size_unit_price)

```

## Code

```{r}
#| eval: false

#abc
```
:::

+---------------------------------------------------------------------------------------------------------------------+
| Insights                                                                                                            |
+=====================================================================================================================+
| -   In terms of size,there seems to be less variability for Executive Condominiums and Apartments and Condominiums. |
|                                                                                                                     |
| -   In terms of unit price per sqft,there seems to be less variability for Executive Condominiums.                  |
+---------------------------------------------------------------------------------------------------------------------+

## EDA 3: Distribution of Transaction counts and unit price across Planning Regions

To visualize the transaction counts among different property types in different Planning regions

We plot a stacked bar chart that shows the transaction counts across the planning regions and the stacks shows the split of counts between different Property Types.

Adding a box plot to understand the variability of price across different Planning regions for each Property Type.

::: panel-tabset
## Plot

```{r, fig.width=18, fig.height=15}
#| echo: false

transaction_counts <- res_data_q1_2024 %>%
  count(Planning_Region,Property_Type) 

p1<- ggplot(transaction_counts, aes(x = Planning_Region, y = n, fill = Property_Type)) +
  geom_bar(stat = "identity", position = "stack") + 
  labs(
    title = "Transaction Count by Property Type and Sale Type (Q1 2024)",
    x = "Property Type",
    y = "Number of Transactions"
  ) +
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 45, hjust = 1)) 



# Create the box plot
p2<-ggplot(res_data_q1_2024, aes(x = Property_Type, y = Unit_Price_sqft)) +
  geom_boxplot() +  facet_grid(~ Planning_Region) +
  labs(
    title = "Price Distribution by Property Type (Q1 2024)",
    x = "Property Type",
    y = "Price"
  ) +
  scale_y_continuous(breaks = seq(0, max(res_data_q1_2024$Unit_Price_sqft), by = 500))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) # Rotate x-axis labels

p1/p2
```

## Code

```{r}
#| eval: false

#abc
```
:::

+------------------------------------------------------------------------------------------------------------------+
| Insights                                                                                                         |
+==================================================================================================================+
| -   Overall sale of apartments seem to be the highest amongst all property types                                 |
|                                                                                                                  |
| -   Apartment sale in Central Region seems to be highest amongst all regions                                     |
|                                                                                                                  |
| -   Box plots shows that higher price outlier data seems to be highest in central area across all Property Types |
|                                                                                                                  |
| -   Box plots shows that higher price outlier data seems to be lowest in North East area for Apartments          |
+------------------------------------------------------------------------------------------------------------------+

## EDA 4: Distribution of students performance

Here we are visualizing the distribution of transaction data across different Sale type ie New Sale, Resale,Subsale in the form of a rain cloud plot.

**Components of a Raincloud Plot:**

-   **"Cloud" (Violin Plot):** This is a rotated and horizontally aligned kernel density estimation that shows the **distribution** of the transaction counts for each planning area. The denser the area of the cloud, the more data points are concentrated at that value.

-   **"Rain" (Strip Plot):** These are individual data points (transaction counts) represented as dots, plotted vertically or jittered slightly to avoid overplotting. They show the **raw data** distribution.

-   **Boxplot:** This summarizes the central tendency (median) and variability (interquartile range) of the transaction counts for each planning area.

::: panel-tabset
## Plot

```{r, fig.width=8, fig.height=6}
#| echo: false
ggplot(res_data_q1_2024, 
       aes(x = Unit_Price_sqft, 
           y = Sale_Type)) +
  #stat_halfeye(adjust = 0.5,justification = -0.3,point_colour = NA) +
  stat_halfeye()+
  #geom_boxplot(width = .20,outlier.shape = NA) + 
  geom_boxplot(width = .15)+
  #facet_grid(~ Property_Type) +
  scale_x_continuous(breaks = seq(0, max(res_data_q1_2024$Unit_Price_sqft), by = 500))+
  #stat_dots(side = "left",justification = 1.3, binwidth = .5,dotsize = 2.5) +
  stat_dots(side = "left")+
 # coord_flip() +
    theme(legend.position = "none", # Hides the legend,
        plot.title = element_text(face = "bold", size = 12),
        plot.subtitle = element_text(size = 8))
```

## Code

```{r}
#| eval: false

#abc
```
:::

+----------------------------------------------------------------------------------------------------+
| Insights                                                                                           |
+====================================================================================================+
| For New sale there are 2 peaks in transaction counts hovering around 1500 and 2250 unit price sqft |
|                                                                                                    |
| There are more higher outlier data in Resale transaction                                           |
+----------------------------------------------------------------------------------------------------+
